---
title: "High flow nasal oxygen during procedural sedation: A randomized controlled trial"
author:
  - Aaron Conway BN(Hons), PhD:
      email: aaron.conway@utoronto.ca
      phone: +1 416 946 7112
      address: 585 University Ave | Toronto, ON  M5G 2N2
      institute: [PMCC, LBFON, QUT]
      correspondence: true
  - Peter Collins MStats:
      institute: [PMCC]
      correspondence: false
  - Kristina Chang BScN, MScN:
      institute: [PMCC]
      correspondence: false
  - Navpreet Kamboj BScN, MScN:
      institute: [LBFON]
      correspondence: false
  - Ana Lopez Filici MHSc(HA),HBSc, RRT, CCAA, FCSRT:
      institute: [TGH]
      correspondence: false
  - Phoebe Lam M.Ed, H.BSc, RRT, CCAA, FCSRT:
      institute: [TGH]
      correspondence: false
  - Matteo Parotto MD, PhD:
      institute: [TGH, UoT]
      correspondence: false
institute:
  - PMCC: Peter Munk Cardiac Centre, University Health Network, Toronto, Canada
  - LBFON: Lawrence S. Bloomberg Faculty of Nursing, University of Toronto, Toronto, Canada
  - QUT: School of Nursing, Queensland University of Technology (QUT), Brisbane, Australia
  - TGH: Department of Anesthesia and Pain Management, UHN, Toronto, Canada
  - UoT: Department of Anesthesia and Interdepartmental Division of Critical Care Medicine, University of Toronto, Toronto, Canada
bibliography: Extras/references.bib
csl: Extras/anesthesia-and-analgesia.csl
output:
  officedown::rdocx_document:
    reference_docx: Extras/manuscript_template.docx
    pandoc_args:
      - --lua-filter=Extras/scholarly-metadata.lua
      - --lua-filter=Extras/author-info-blocks.lua
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(magrittr)
library(patchwork)
library(drake)

# Values of interest
loadd(CI_co2_peak)
loadd(CI_co2_mean)
loadd(CI_isas)
loadd(CI_diffoxygen)
loadd(CI_diffoxygen_best)
loadd(CI_diffoxygen_worst)
loadd(CI_diffuseoxygen)
loadd(CI_diffuseoxygen_best)
loadd(CI_diffuseoxygen_worst)
loadd(CI_troops)
loadd(CI_spo2)

loadd(comfortprob)
loadd(aucprob)
loadd(desatprob)
loadd(troopsprob)
```


**Acknowedgements:** This study was funded by a Connaught New Researcher Award (University of Toronto). MP is supported by an Early Investigator Merit Award from the Department of Anesthesia, University of Toronto and Toronto General Hospital.

**Conflicts of interest:** The authors declare no competing interests.

**Keywords:** Anesthesia, oxygen supplementation, deep sedation, conscious sedation, cardiology.

\pagebreak


### Abstract

**Objective**

High flow nasal oxygen (HFNO) may better support the vulnerable breathing state of patients during procedural sedation. The objective of this study was to investigate the effects of HFNO in comparison to facemask oxygen during cardiac implantable electronic device (CIED) procedures performed with procedural sedation.

**Methods**
A randomized controlled trial was conducted. Participants were 1:1 randomized to facemask (≥ 8L/min) or HFNO (50L/min with a 50:50 oxygen to air ratio) during procedural sedation administered by Anesthesia Assistants (supervised by an Anesthesiologist). The primary outcome was peak transcutaneous carbon dioxide (TcCO~2~). Outcomes were analysed using Bayesian statistical models.

**Results**

The 129 participants who were randomized and received a CIED with sedation were included. The difference in peak TcCO~2~ was `r CI_co2_peak[["fullCI"]]`. Minor adverse sedation events were `r CI_troops[["effect"]]` times more likely to occur in the HFNO group. This estimate is imprecise `r CI_troops[["CIbrackets"]]`. The odds ratio for oxygen desaturation for the HFNO group compared with the facemask group was `r CI_spo2[["effect"]]` `r CI_spo2[["CIbrackets"]]`. The difference in satisfaction with sedation scores betwen groups was `r CI_isas[["fullCI"]]`.

**Conclusions**

We investigated the effects of using HFNO with oxygen to air ratio settings estimated to deliver a FiO2 approximately equivalent to that achieved from standard practice with facemask oxygen. There was no clear advantage for using HFNO in preference to facemask oxygen during CIED procedures performed with sedation.

**Trial registration number:** NCT03858257

**Section head:** Healthcare delivery, economics and global health

**What is already known about this subject?**

- A prior physiological modeling study identified a mechanism by which HFNO promotes carbon dioxide clearance.

**What does this study add?**

- We found that HFNO at 50L/min for patients undergoing elective CIED procedures with sedation is highly unlikely to *decrease* or *increase* peak TcCO~2~ concentration by a clinically important amount.

**How might this impact on clinical practice?**

- Using HFNO during CIED procedures performed with procedural sedation is not likely to prevent potential adverse sedation events that may be precipitated by hypercarbia.

\pagebreak

## Introduction

Cardiac implantable electronic device (CIED) procedures are commonly performed with procedural sedation.[@conway2014trends] Oxygen supplementation is administered to reduce hypoxemia from sedation-induced hypoventilation.[@Conway_2019; @conway2013risk] High flow nasal oxygen (HFNO) is a promising device for oxygen supplementation.[@drakeHighFlowNasalCannula2018]  HFNO allows for heated, humidified gas with a titratable oxygen:air ratio to be administered via nasal prongs at up to 70L/min. Delivering oxygen supplementation at such high flow-rates has physiological effects that may support the vulnerable breathing state of patients during procedural sedation. In particular, one of the proposed physiological effects of HFNO is that it facilitates active gas exchange during times of apnea due to turbulent supraglottic flow vortices.[@Hermez_2019] The effects of the potential disadvantages of using HFNO during sedation should also be evaluated. It is possible that the potential gains arising from the HFNO device may be offset by the reduced ability to monitor ventilation from capnography waveforms when it is being used, as exhaled carbon dioxide concentrations are “washed out” by the high gas flow. Guidelines from the American Society of Anesthesiology have stated that there is insufficient evidence regarding which supplemental oxygen device (e.g., nasal cannula, face mask, or specialized devices such as HFNO) is most effective.[@apfelbaum2018practice] The objective of this study was to investigate the effects of HFNO in comparison to facemask oxygen during CIED procedures performed with procedural sedation.

## Methods

### Design

A randomized controlled trial design was used with participants 1:1 randomized to:

1. Facemask oxygen; or
2. High flow nasal oxygen.

Informed consent was obtained. The study protocol conforms to the 1975 Declaration of Helsinki and received approval by the hospital research ethics board (18-6343). The trial was prospectively registered (NCT03858257).

### Participants

Adults undergoing an elective CIED procedure with sedation administered by an Anesthesia Assistant at one large academic teaching hospital in Canada were included.

*Exclusion criteria*

1. <16 years. 
2. Underlying condition requiring chronic oxygen supplementation.
3. Diagnosed respiratory condition with current hypercapnia defined as PaCO~2~ during admission over 45mmHg.
4. Pre-existing untreated pneumothorax.
5. Planned transesophageal echocardiography.
6. Active nasal-bleeding.
7. Complete nasal obstruction.
8. Recent upper-airway surgery or base of skull fracture.
9. Previous participation.

### Interventions

**Facemask oxygen supplementation**

Supplemental oxygen through a facemask with the flow-rate chosen by the Anaesthesia Assistant as per their standard practice.

**High flow nasal oxygen**

The Optiflow device (Fisher and Paykel Healthcare, Auckland, New Zealand), heated breathing tube and chamber, and nasal cannula was used. This system is a humidifier with an integrated flow generator, able to humidify respiratory gases and deliver them down a heated breathing tube and through the nasal cannula interface. The gas temperature was set to the 'High' setting (ranges 30-32º Celsius) and titrated downwards if the patient complained of irritation. The gas flow-rate was commenced at 30L/min prior to sedation administration and titrated up to 50L/min as tolerated by the patient after sedative medication was administered. The fraction of oxygen in the gas was commenced at 50% but could be titrated according to patient requirements.

**Concomitant care**

There were no restrictions on concomitant care. Anesthesia Assistants were permitted to use standard physioligcal monitoring devices, as dictated by the Canadian Anesthesiologists' Society (CAS), and to titrate sedation according to their usual practice.[@Dobson_2018] Concomitant care most relevant to this trial was the use of capnography. Anesthesia Assistants elected to use capnography regardless of whether supplemental oxygen was delivered via HFNO or facemask, as this is a requirement from the Canadian Anesthesiologist Society anytime procedural sedation is being administered.[@Dobson_2018] The facemask had an integrated CO~2~ sampling line. For participants randomized to HFNO, Anesthesia Assistants used the CO~2~ sampling adapter integrated with the latest model of the HFNO nasal cannula for the majority of participants (all those recruited after September 2019). Prior to this model becoming available, Anesthesia Assistants placed a facemask with an integrated CO~2~ sampling line over the HFNO nasal cannula. Oxygen supplementation was delivered through the HFNO nasal cannula and CO~2~ was sampled from the sampling line integrated into the facemask.


### Outcomes

Outcome selection was informed by recommendations from the Sedation Consortium on Endpoints and Procedures for Treatment, Education and Research (SCEPTER).[@Ward_2018] The primary outcome was peak transcutaneous carbon dioxide (TcCO~2~) concentration. Secondary outcomes were:

1. Mean TcCO~2~.
2. Trajectory of TcCO~2~ as a function of time.
3. Area under the curve of oxygen desaturation (AUC~DESAT~). This is the difference between the threshold (90%) and actual oxygen saturation (SpO~2~) summed every minute during which oxygen saturation was below the threshold.
4. Adverse sedation events, measured using the Tracking and reporting outcomes of procedural sedation (TROOPS) tool.
5. Patient satisfaction with sedation.
7. Comfort of the oxygen delivery device.
8. Anesthesia Assistant rating of difficulty maintaining oxygenation status.
9. Anesthesia Assistant rating of difficulty using oxygen delivery device.


### Data collection

#### Instruments

TcCO~2~ was measured continuously using the Sentec Digital Monitoring system with VSign 2 sensor. TcCO~2~ monitoring provides continuous, accurate and precise estimates of PaCO~2~.[@conway2019accuracy] TcCO~2~ monitoring may provide even more precise estimates of changes in PaCO~2~ (mean bias 0.03 mmHg, 95% limits of agreement -0.44 to 0.38 mmHg).[@Fingleton2017] The Sentec VSign 2 sensor was attached to the forehead. Once the TcCO~2~ stabilized, the monitor was covered with a drape so that it was not visible to research staff or clinicians. The monitor was not used by the clinicians to guide treatment. TcCO~2~ was sampled at a frequency of one measurement per second. The recorded SpO~2~ was extracted from the Drug Reconciliation and Electronic Monitoring System at a frequency of one measurement per minute. Adverse sedation events were measured using the tracking and reporting outcomes of procedural sedation (TROOPS) tool.[@Roback2018] Satisfaction with sedation was measured using the Iowa Satisfaction with Anesthesia Scale (ISAS).[@dexter1997development] Participants were asked to rate comfort with the oxygen delivery device and Anesthesia Assistants were asked to rate their: 1) perceived level of difficulty in maintaining oxygenation; and 2) perceived level of difficulty in using the oxygen delivery device, using a 6-level rating scale.


### Sample size calculation

We estimated based on our prior work[@Conway_2019] that the peak TcCO~2~ level in the control group would be 47 mmHg and standard deviation would be 7 mmHg. Assuming a type I error rate of 5%, a sample of 130 participants would achieve 90% power to detect a reduction in mean TcCO~2~ levels of 4 mmHg in the intervention period. A difference in TcCO~2~ levels of 4 mmHg was selected for this sample size calculation because it was deemed of potential clinical relevance and was used to power previous trials.[@Wijesinghe2012]  Differences in CO~2~ level of a similar magnitude have been detected in previous trials evaluating the efficacy of interventions to improve sedation safety.[@Baulig2015; @Smith2015] 

### Random sequence generation and concealment

A stratified (by diagnosis of obstructive sleep apnea and type of procedure – cardiac resynchronization therapy device implant), block randomized sequence was generated and concealed using the web-based randomization feature in REDCap^TM^. The RA retrieved the allocation for each consecutive participant in REDCap^TM^ prior to the procedure.


### Statistical analyses

Bayesian statistical models were used. Data and code are available [here](https://hfnosedrct.netlify.app/flexdashboard) and is archived [here](https://doi.org/10.5281/zenodo.3908492).  A detailed summary of the statistical models is presented in the [Appendix](https://hfnosedrct.netlify.app/appendix). Prior distributions were chosen to be weakly informative, which is appropriate in the absence of information concerning the likely values of model parameters.[@gelman2008weakly] Covariate adjustments were made for the stratification variables obstructive sleep apnea (OSA) status and whether or not the procedure was a cardiac resynchronization therapy (CRT) device implant as well as for baseline TcCO~2~ concentration, which was modelled using splines.[@Egbewale2014] Continuous outcomes were analyzed using robust regression models. A functional analysis of variance (ANOVA) model was used to investigate how mean TcCO~2~ concentration levels differ between groups as a function of procedure time [@yue2019]. Logistic regression was used for dichotomous outcomes. Proportional-odds models were used for ordinal outcomes. Analysis was performed only on those participants whose SPO~2~ was observed to fall below the 90% threshold for the AUC~DESAT~ outcome.

Posterior inference for all models except the functional ANOVA model was performed using Hamiltonian Monte Carlo through the brms package [@bruckner2017], version `r packageVersion("brms")`. For this set of models, 2000 posterior samples were obtained from 4 independent chains of 2000 samples, where the first 1000 warm-up samples were discarded. Posterior inference for the functional ANOVA model was performed using the Integrated Nested Laplacian Approximation [@rue2009] through the INLA package, version `r packageVersion("INLA")`. The marginal posterior distribution of parameters was summarized by their mean and a 95% credible interval defined by the interval spanning the 2.5% and 97.5% percentiles of their distributions. The clinical significance of treatment effects relating to TcCO~2~ concentration were evaluated by computing the posterior probability that an effect exceeds 4 mmHg in either direction. When the proportion of missing data was large and the missing completely at random (MCAR) assumption was unlikely to be satisfied, a sensitivity analysis was performed to investigate the robustness of the conclusions of the complete-case analysis.


### Patient and public involvement

This article was written without patient involvement. Patients were not invited to comment on the study design and were not consulted to develop patient-relevant outcomes or to interpret the results. Patients were not invited to contribute to the writing or editing of this document for readability or accuracy.

## Results


### Participants

From August 2019 to March 2020, we screened 270 patients undergoing CIED procedures (Figure 1). A total of 130 participants were randomized. One participant was excluded because the procedure was cancelled. One participant, who was randomized to the HFNO group, had their procedure rescheduled to a time that the Research Assistant was not available. As such, this participant received oxygen via standard face mask and TcCO~2~ data were not collected. For two participants, the TcCO~2~ sensor failed to calibrate prior to commencement of the procedure. Most (n=29; 45%) Anesthesia Assistants reported having used HFNO between 2-5 times.

Participant characteristics are presented in Table 1. The sample was mostly elder and male. Anesthesia Assistants' rated the ASA Physical Classification Status as either III or IV. Obstructive sleep apnea was common. About 20% of procedures were for cardiac resynchronisation therapy. 

### Comparisons between groups

#### Primary outcome

Results are presented in Table 2. The effect of HFNO on the peak TcCO~2~ was estimated to be `r CI_co2_peak[["fullCI"]]`. The probability that it exceeds the 4mmHg clinical significance threshold of 4mmHg in either direction is 0.

#### Secondary outcomes

The effect of  HFNO on the mean TcCO~2~ concentration was estimated to be `r CI_co2_mean[["full"]]`. The probability that it exceeds the 4mmHg clinical significance threshold is 0 in either direction.  TcCO~2~ concentrations for all patients throughout procedures are displayed in Figure 2, with the longest procedure highlighted as a reference. The estimated effect did not exceed the 4mmHg clinical significance threshold  in either direction with probability greater than 0.95. There is no discernable trend observed in how the effect varies with procedure time. Precision decreases as time increases, reflecting the shrinking number of participants.

The effect of HFNO on ISAS score was estimated to be `r CI_isas[["fullCI"]]`. The probability that patients are more likely to rate comfort with the oxygen supplementation device higher with HFNO compared to the facemask is `r comfortprob$above`.
 
The odds ratio for Anesthesia Assistant ratings of difficulty maintaining oxygenation status and difficulty using the oxygen delivery device as estimated using a complete-case analysis are `r CI_diffoxygen[["fullCI"]]` and `r CI_diffuseoxygen[["fullCI"]]`, where a value less than 1 indicates a greater level of difficulty for respondents in the HFNO group. It should be noted, however, that the Anesthesia Assistants' ratings of difficulty using the oxygen device and difficulty maintaining oxygenation were missing 45 and 46 responses, respectively, likely due to the survey being voluntary. It is unlikely that missingness among these ratings occurred completely at random, so a best- and worst-case imputation approach was used to investigate the impact that the missing data could have on the results in extreme cases. The best- and worst-case sensitivity analysis gave estimates ranging between `r CI_diffoxygen_worst[["fullCI"]]` and `r CI_diffoxygen_best[["fullCI"]]` for difficulty maintaining oxygenation status and from `r CI_diffuseoxygen_worst[["fullCI"]]` and `r CI_diffuseoxygen_best[["fullCI"]]` for difficulty using the oxygen delivery device.

The odds ratio for a minor adverse sedation event related to airway or breathing for the HFNO group compared with the facemask group was `r CI_troops[["effect"]]`. This effect estimate is very imprecise due to the small number of events (95% CI `r round(CI_troops[["lower"]], 1)` to `r round(CI_troops[["upper"]],0)`). A similar number of participants in the HFNO group (n=8; 12%) experienced an oxygen desaturation event in comparison with the facemask oxygen group (n=7; 11%). `r CI_spo2[["fullCIbrackets"]]`. The effect estimate for the absolute difference in the AUC~DESAT~ was imprecise, spanning from 5 minutes.% higher in the face mask group to 24 minutes.% higher in the HFNO group. The probability that AUC~DESAT~ is higher with HFNO is `r aucprob$above`.  A visualization of the SpO~2~ trajectories for patients whose SpO~2~ was below 90% is available [here](https://hfnosedrct.netlify.app/flexdashboard). 

#### Oxygen flow-rates

Most participants randomized to the HFNO group had the flow-rate set at 50L/min (Figure 3). Most participants randomized to the facemask group received oxygen at ≥8L/min. Two participants who were randomized to HFNO did not receive this intervention at all and four participants who were randomized to HFNO stopped receiving this intervention at a certain timepoint during procedures at the discretion of the Anesthesia Assistant, with the rationale that capnography monitoring was not sufficient with the HFNO device. 

## Discussion

We found that HFNO at 50L/min for patients undergoing elective CIED procedures with sedation is highly unlikely to *decrease* or *increase* peak TcCO~2~ concentration by a clinically important amount. Although a prior physiological modeling study identified a mechanism by which HFNO promotes carbon dioxide clearance,[@Hermez_2019] it seems the magnitude of any such effect is insufficient to produce an important difference in ventilation status during procedural sedation. Importantly, we used continuous TcCO~2~ monitoring so that we could estimate differences in ventilation between groups over the whole duration of procedures. There was no discernible trend observed in how the effect varied over time, lending strength to the assertion that ventilation status is not impacted by the use of HFNO.

The probability that minor adverse sedation events related to airway and breathing are more likely to occur with HFNO is `r troopsprob$above`. The suspected etiology noted for these events by the Anesthesia Assistants in the TROOPS tool was oxygen desaturation. There are two plausible mechanisms that may explain this result. It is possible that the oxygen:air blend (50:50) used in the HFNO group was simply not equivalent to the amount of oxygen supplementation received in the facemask group. Most participants in the facemask group received >8L/min of 100% O~2~.  Further research with a larger sample size would be useful to determine the optimal oxygen:air ratio for HFNO during sedation for CIED procedures, with a focus on adverse sedation events or hypoxemia as the primary outcome.

Another plausible mechanism is that the ability to monitor capnography waveforms was diminished with HFNO.  Capnography is widely considered to be an essential aspect of physiological monitoring during sedation.[@conway2016systematic; @conway2019pre; @conway2014clinical] The concern about reduced ability to monitor capnography waveforms when HFNO is used potentially increasing the risk of more prolonged, undetected episodes of hypoventilation during sedation has been raised previously in the literaure.[@horiuchi2020high] However, it should be noted that if undetected episodes of hypoventilation were considerably more frequent and prolonged when HFNO was used in our study, presumably, we would have observed higher TcCO~2~ concentrations in this group. We did not observe higher TcCO~2~ concentrations in the HFNO group for the peak measurement or at any particular time-point during procedures. 

The evidence base for the effects of HFNO therapy for procedural sedation in other clinical contexts is limited. One large[@lin2019high] and three small randomized controlled trials were published in 2019, with several more on-going trials registered.[@eugene2020high] The primary outcomes for all the trials to date have focused on investigating the impact of HFNO on oxygenation with inconsistent results. One of the small trials randomized 60 participants undergoing bronchoscopy to receive HFNO at 50L/min with 100% oxygen or oxygen at 10-15L/min through a facemask.[@douglasRandomisedControlledTrial2018]  There was no difference observed between the treatment groups for the primary outcome, which was the proportion of patients who experienced oxygen desaturation (defined as SpO~2~ 90%). Another trial randomized 59 morbidly obese patients undergoing endoscopy to receive a fraction of inspired oxygen concentration of 0.36 either via HFNO at a flow-rate of 60L/min or via nasal cannula at 4L/min.[@riccio2019high]  Again, there was no difference in the primary outcome of oxygen desaturation (SpO~2~ < 90%). The third study randomized 30 participants undergoing dental sedation into three groups to receive a fraction of inspired oxygen concentration of 0.4 either via HFNO at a flow-rate of 50L/min, via HFNO at a flow-rate of 30L/min or via nasal cannula at 5L/min.[@sago2015nasal]  Participants randomized to the HFNO groups had higher nadir blood oxygen levels recorded than the low flow oxygen group. In contrast, a large trial of 1994 participants undergoing gastroscopy with propofol sedation reported a large reduction in risk of hypoxemia (8.4% in the control group and 0% in the HFNO group).[@lin2019high] This result is likely explained by the large difference in FiO~2~ that was delivered between the two groups. In the HFNO group participants received 60L/min of 100% oxygen and in the control group participants received just 2L/min of oxygen.  

Satisfaction with sedation is very likely to be similar between HFNO and facemask oxygen. The probability that patients are more likely to rate comfort with the oxygen supplementation device higher with HFNO was `r comfortprob$above`. In contrast, we identified that the HFNO device was rated as more difficult for Anesthesia Assistants to use compared with the standard facemask. None of the Anesthesia Assistants rated the HFNO device as *difficult* to use and most had very limited experience using the device. Also, most Anesthesia Assistant participants reported they had used HFNO between 2 and 5 times. Experience with HFNO is likely to influence clinicians' perceptions about the difficulty using the device.


### Limitations

The primary outcome was peak TcCO~2~ and we accounted for the correlation between baseline and peak measurements by including the baseline measurements as a covariate in the model. However, a potential limitation is that results may be sensitive to how the baseline and peak measurements were chosen. We did not blind participants or clinicians to group assignment. The small dropout and cross-over rate are unlikely to have exerted a major impact on the effect estimates. Participants received propofol, midazolam and fentanyl, which is a common and recommended approach for CIED procedures.[@furniss2015safe] Severe oxygen desaturation is not a common event when oxygen supplementation is delivered at flow-rates between 6-10L/min through a face mask during procedures performed with sedation.[@conway2013risk; @conway2014trends]  Results from our trial cannot be directly generalized to other clinical settings where desaturation is more severe and occurs more often.

### Conclusion

We investigated the effects of using HFNO with the flow-rate and oxygen to air ratio set to deliver a FiO~2~ approximately equivalent to that achieved from standard practice with facemask oxygen. Ventilation, as measured by TcCO~2~, is highly unlikely to differ by a clinically important amount. The probability that minor adverse sedation events were more likely to occur in the HFNO group was high and the severity of oxygen desaturations is probably worse with HFNO. There is a higher probability that patients will be more comfortable during procedures with HFNO in comparison to the facemask, but overall patient satisfaction with sedation is likely to be similar. Overall, there was no clear advantage identified for using HFNO in place of facemask oxygen.

\pagebreak


# References

<div id="refs"></div>

\pagebreak

# Figure legend
    
Fig. 1 CONSORT Flow Diagram

Fig. 2 Transcutaneous carbon dioxide measurements throughout procedures

Fig. 3 Oxygen flow-rates

\pagebreak


## Table 1. Participant characteristics


```{r }
drake::readd(characteristics_table)
```

<!---BLOCK_MULTICOL_START--->


<!---BLOCK_STOP--->


<!--BLOCK_LANDSCAPE_START--->

## Table 2. Results

```{r }
readd(flextable_outcome)
```


<!---BLOCK_LANDSCAPE_STOP--->

